---
title: "R Notebook"
output: html_notebook
---

# Comparative Genomics

```{r}
setwd("C:/Users/abish/Documents/Online Learning/R Learning/Bioconductor")
```

 install the biomaRt package from Bioconductor
```{r}
BiocManager::install('biomaRt')
```

Biomart lets you access ENSEMBLE genomes directly and download relevant information, we can see a list of available databases (marts) using this command
```{r}
library(biomaRt)
listMarts()
```

```{r}
ensembl=useMart("ensembl")
```

```{r}
listDatasets(ensembl)
```

load the data for the Chimpanzee and the Gorilla datasets
```{r}
Chimpanzee <- useDataset("ptroglodytes_gene_ensembl", mart = ensembl)

Gorilla <- useDataset("ggorilla_gene_ensembl", mart = ensembl)
```

```{r}
ChimpanzeeAttributes <- listAttributes(Chimpanzee)

GorillaAttributes <- listAttributes(Gorilla)
```

```{r}
GorillaAttributes
```

The ensemble Gene ID is a unique gene identifier and can be used to indicate a specific gene regardless of which species it is in – let's use that in our analysis.
```{r}
ChimpanzeeGenes <- getBM(attributes = c("ensembl_gene_id", "gene_biotype"), mart=Chimpanzee)

GorillaGenes <- getBM(attributes = c("ensembl_gene_id", "gene_biotype"), mart=Gorilla)
```

parse the gene names and gene types into two separate vectors for each species
```{r}
ChimpanzeeGeneNames <- ChimpanzeeGenes[[1]]

ChimpanzeeGeneTypes <- ChimpanzeeGenes[[2]]

GorillaGeneNames <- GorillaGenes[[1]]

GorillaGeneTypes <- GorillaGenes[[2]]
```

```{r}
ChimpanzeeTypesTable <- table(ChimpanzeeGeneTypes)

GorillaTypesTable <- table(GorillaGeneTypes)
```

```{r}
ChimpanzeeTypesTable
```

compare the number of protein coding genes in gorilla vs chimpanzees.
```{r}
ChimpanzeeProteins <- ChimpanzeeTypesTable["protein_coding"]

GorillaProteins <- GorillaTypesTable["protein_coding"]
```

Let's see if we can identify which genes Chimpanzees and Gorillas have in common.

One of the attributes in the Chimpanzee dataset is called "ggorilla_homolog_ensembl_gene" this is the gene identifier for the gorilla gene that is orthologous to the chimpanzee gene.

First let's just see how many genes Chimpanzees have that have gorilla orthologs

```{r}
ChimpGGNum <- getBM(attributes = "ggorilla_homolog_ensembl_gene", mart = Chimpanzee)
```

```{r}
ChimpanzeeGG <- getBM(attributes = c("ensembl_gene_id","ggorilla_homolog_ensembl_gene"), mart = Chimpanzee)
```

```{r}
GorillaCNum <- getBM(attributes = "ptroglodytes_homolog_ensembl_gene", mart = Gorilla)

GorillaC <- getBM(attributes = c("ensembl_gene_id","ptroglodytes_homolog_ensembl_gene"), mart = Gorilla)
```

The GorillaCNum tells us that gorillas have almost 27k genes in common with Chimpanzees
The GorillaC table lists them by ID
