---
title: "R Notebook"
output: html_notebook
---

RNA Sequencing

```{r}
BiocManager::install("edgeR")
```

```{r}
BiocManager::install("Glimma")
```
```{r}
library(edgeR)

library(limma)

library(Glimma)

library(org.Mm.eg.db)
```

```{r}
library(gplots)

library(RColorBrewer)
```

gene count information for each sample
```{r}
seqdata <- read.delim("GSE60450_Lactation-GenewiseCounts.txt", stringsAsFactors = FALSE)

```
metadata about our samples
```{r}
sampleinfo <- read.delim("SampleInfo.txt")
```

make a new dataset that contains only the counts, for manipulation – we need to remove the first two columns from seqdata using this command
```{r}
countdata <- seqdata[,-(1:2)]
```
```{r}
rownames(countdata) <- seqdata[,1]
```

use the substr command to extract just a portion of the characters from each of the
column names in seqdata
```{r}
colnames(countdata) <- substr(colnames(countdata),start=1,stop=7)
```

extract the CPMs (counts per million reads) and examining them
```{r}
myCPM <- cpm(countdata)
```

filter out anything with a CPM of less than 0.5
```{r}
thresh <- myCPM > 0.5
```

quick statistics to see how many genes passed through our filter
```{r}
table(rowSums(thresh))
```

keep only those where the genes are TRUE in 2 or more samples
```{r}
keep <- rowSums(thresh) >= 2

counts.keep <- countdata[keep,]
```

```{r}
plot(myCPM[,1],countdata[,1])
```
```{r}
plot(myCPM[,1],countdata[,1],ylim=c(0,50),xlim=c(0,3))
```

